---
title: "Initial exploration"
author: "Diabb Zegpi D."
date: "11-12-2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(here)
theme_set(theme_minimal())

col_spec <- cols(
  VAR6_sum = col_double(),
  VAR6_prom = col_double(),
  VAR6_trx = col_double(),
  VAR9_sum = col_double(),
  VAR9_prom = col_double(),
  VAR9_trx = col_double(),
  VAR30_sum = col_double()
)

bci_train <- read_csv(here("Data", "train_data.csv"), col_types = col_spec)
bci_test <- read_csv(here("Data", "test_data.csv"), col_types = col_spec)
```


## Training set

- Total rows = 667691
- Total cols = 70
- Response = target_mes
- VARx_sum = total mes
- VARx_prom = promedio diario del mes
- VARx_trx = total operaciones por mes
- VARx_ratio = índice promedio por mes
- There is a ton of missing data, due to lack of information of clients operations and lack of transactions.
- From the mes column can be extracted month and year.

```{r}
bci_train %>% 
  mutate(month = as.numeric(str_sub(mes, start = -2L)),
         year = as.numeric(str_sub(mes, end = -3L)))
```


```{r}
percentage <- function(x, ...) {
  x %>% 
    count(..., sort = TRUE) %>% 
    mutate(pct = n / sum(n),
           cum_pct = cumsum(pct))
}

bci_train %>% percentage(tipo_ban) # Class imabalance towards BAN1 (83%)
bci_train %>% percentage(tipo_seg) # Maybe recode this col
bci_train %>% percentage(categoria) # Few examples of V1
bci_train %>% percentage(tipo_com) # All weel representated
bci_train %>% percentage(tipo_cat) %>% print(n = Inf) # Maybe keep top 90%
```


```{r}
bci_train %>% 
  select(starts_with("VAR")) %>% 
  summarise(across(.fns = function(x) sum(is.na(x)) / nrow(bci_train))) %>% 
  pivot_longer(cols = everything()) %>% 
  mutate(name = fct_reorder(name, -value)) %>% 
  ggplot(aes(name, value)) +
  geom_col(alpha = .8, fill = "deepskyblue4", color = "white") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Percentage of NA for each numeric variable",
       x = NULL, y = "% of missing rows") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, margin = margin(-20)),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank())
```

All the numeric VARx like columns present, at least, 20% of missingness. 


```{r}
bci_train %>% 
  select(contains("sum")) %>% 
  pivot_longer(cols = everything()) %>%
  mutate(name = fct_inorder(name),
         value = value + 1) %>% 
  na.omit() %>% 
  ggplot(aes(value, ..density..)) +
  geom_histogram(bins = 30) +
  scale_x_log10() +
  facet_wrap(~name, ncol = 6)


bci_train %>% 
  select(contains("prom")) %>% 
  pivot_longer(cols = everything()) %>%
  mutate(name = fct_inorder(name),
         value = value + 1) %>% 
  na.omit() %>% 
  ggplot(aes(value, ..density..)) +
  geom_histogram(bins = 30) +
  scale_x_log10() +
  facet_wrap(~name, ncol = 6)


bci_train %>% 
  select(contains("trx")) %>% 
  pivot_longer(cols = everything()) %>%
  mutate(name = fct_inorder(name),
         value = value + .1) %>% 
  na.omit() %>% 
  ggplot(aes(value, ..density..)) +
  geom_histogram(bins = 30) +
  scale_x_log10() +
  facet_wrap(~name, ncol = 6)


bci_train %>% 
  select(contains("ratio")) %>% 
  na.omit() %>%
  mutate(VAR29_ratio = VAR29_ratio + .001) %>% 
  ggplot(aes(VAR29_ratio, ..density..)) +
  geom_histogram(bins = 30) +
  scale_x_log10() 
```


```{r}
# Variables with too many zeros (sum)
# "VAR14_sum", "VAR15_sum", "VAR18_sum", "VAR19_sum", 
# "VAR20_sum", "VAR21_sum", "VAR22_sum"
# prom variables don't have problems with zeros
# There are many trx variables with a ton of zeros. A dummy can be made.
# VAR29_ratio has little zeros but many missing values.
# All numeric variables should be transformed with the logarithm

library(tidymodels)
tidymodels_prefer()

set.seed(123)
bci_split <- initial_split(bci_train)
train <- training(bci_split)
test <- testing(bci_split)

exclude_vars <- c("VAR14_sum", "VAR15_sum", "VAR18_sum", "VAR19_sum", 
                  "VAR20_sum", "VAR21_sum", "VAR22_sum")

lm_rec <- recipe(target_mes ~ ., data = train) %>% 
  step_mutate(month = as.numeric(str_sub(mes, start = -2L)),
              year = as.numeric(str_sub(mes, end = -3L)),
              across(starts_with("VAR"), function(x) {
                if_else(is.na(x), 0, x)
              })) %>% 
  step_rm(mes, all_of(exclude_vars)) %>% 
  step_log(contains("sum"), offset = 1) %>% 
  step_log(contains("prom"), offset = 1) %>% 
  step_log(contains("trx"), offset = .01) %>% 
  step_other(tipo_cat, threshold = .05, other = "RUBRO_otro") 
  # step_dummy(all_nominal())



lm_rec_prep <- lm_rec %>% prep()


step_dummy(all_nominal()) %>% prep()

```


```{r}

```



















